\section{Conclusion}
\label{sec:conclusion}

This work addresses the optimization of contractual coverage for medical equipment, with the operational goals of reducing the fraction of uncovered equipment and improving contractual decision-making. The problem is formalized as a finite-horizon Markov decision process on a single piece of equipment and solved with the PPO algorithm, trained entirely in a reproducible simulator built in the absence of direct access to Philips France's operational data. The contribution is applicative: we do not propose a new algorithm, but we provide a clean MDP formulation, a reproducible training pipeline, and a statistically grounded comparison with three baselines: uniform random actions, a threshold heuristic whose parameter is swept, and a greedy policy from discounted value iteration on the tabulated expected MDP.

On $100$ evaluation episodes drawn from a shared seed pool, the PPO agent reaches a mean reward of $54.56 \pm 3.93$ ($95\%$ confidence interval $[53.79, 55.33]$), far above the random policy ($-241.44 \pm 92.66$) yet still below both the heuristic ($58.06 \pm 3.93$) and VI ($58.08 \pm 3.96$), with Welch tests against PPO at $p < 10^{-4}$ for either co-optimal reference, while heuristic and VI means are statistically indistinguishable ($p \approx 0.97$). Threshold sweeps trace a flat peak at $58.06$ before decaying to $49.14$ at threshold $15$: PPO therefore sits between the loosest heuristic and the Bellman-aligned plateau.

A state-by-state interpretability tabulation localises the disagreement: PPO matches VI on $95.2\%$ of $65{,}880$ reachable states, with residual mistakes concentrated in covered-state \emph{do nothing} versus \emph{extend} trade-offs. Taken together, these observations show that PPO recovers non-trivial behaviour from reward alone but does not match any certificate of Bellman optimality available in this toy model. Reinforcement learning's comparative advantage is expected only once portfolios, partial observability and budget couplings break the parsimony that tabulation already exhausts here.

These results rely on a deliberately simplified simulator, a four-dimensional state representation and a training budget of $5{,}000$ episodes, all of which bound the scope of the conclusions. Natural extensions include the exploitation of real operational data to enrich the state representation, the extension of the framework to heterogeneous portfolios subject to global budget constraints, a sensitivity study of PPO itself to changes in reward weights and environmental parameters, and ultimately a supervised deployment in production.

The presented approach demonstrates the viability of driving contractual decisions through reinforcement learning and opens the way to a gradual, empirically justified adoption of learned policies complementary to existing business rules.