\section{Introduction}
\label{sec:introduction}

In large healthcare organizations such as Philips France, the management of medical equipment relies on a complex portfolio of maintenance contracts, manufacturer warranties and warranty extensions. This management is critical: it directly conditions the continuity of clinical services, the quality of patient care, and the operational costs borne by the organization. The contractual portfolio evolves continuously, driven by installations, renewals and expirations, which makes coverage mastery particularly difficult at the scale of a large installed base.

Despite the growing availability of data from information systems, existing tools remain largely descriptive and fragmented. Current dashboards offer partial visibility into contractual coverage, but they neither anticipate risks associated with contract expirations nor prioritize actions in a proactive manner. They also fail to provide explanations accessible to business users, limiting their ability to interpret indicators, justify decisions, and act quickly in complex operational environments. The result is a growing need for intelligent solutions capable of transforming data into actionable decisions while delivering clear, contextualized justifications.

We argue that this problem is naturally expressed as sequential decision-making under uncertainty and is therefore well suited to a reinforcement learning (RL) treatment. Specifically, we formalize it as a \emph{Markov decision process} (MDP) in which the state describes the set of managed pieces of equipment and their current contractual posture (maintenance, manufacturer warranty, extension), the actions correspond to coverage decisions (subscription, renewal, adjustment, prioritization), transitions model the portfolio dynamics and contract expirations, and the reward balances coverage cost against the penalty associated with uncovered equipment and late interventions. In this framework, an RL agent learns a long-horizon coverage policy that anticipates expirations and prioritizes the actions with the greatest value for the organization.

Operationalizing this formulation raises several challenges. The environment is only partially observable and must be reconstructed from heterogeneous data sources; the action space combines discrete decisions over a large number of pieces of equipment; the cost of erroneous exploration in production precludes direct online learning, which motivates the use of simulators and offline evaluation protocols; finally, the learned policy must be interpretable to be adopted by the business teams responsible for contract management.

This work is positioned as an \emph{applied case study}. We do not claim a methodological contribution: the PPO algorithm used here is standard and has been extensively studied in prior work~\cite{schulman2017ppo, andrychowicz2021matters}. Our contribution lies in the problem formulation, the reproducible experimental setup, and the rigorous head-to-head comparison with business-rule baselines in a domain---the coverage management of hospital equipment---that, to our knowledge, has not been cast as a reinforcement-learning problem in the published literature.

More specifically, the contributions of this work are:
\begin{itemize}
    \item \textbf{Problem formalization.} We formalize single-equipment contract-coverage optimization as a finite-horizon Markov decision process, with a four-dimensional state aggregating coverage status, time to expiration, contract cost level and risk level, and a three-way action space covering the standard contractual operations (do nothing, renew, extend).
    \item \textbf{Reproducible simulator.} We develop a self-contained simulation environment reproducing the coverage, expiration and intervention dynamics under parametric control, with fully deterministic reproducibility via seeded initial states and failure processes.
    \item \textbf{Bellman reference.} We compute a greedy tabular optimum through discounted value iteration on an augmented finite MDP that mirrors the simulator's transition logic in expectation. This certificate clarifies achievable reward under our modelling commitments and highlights when rule-based controllers are analytically aligned with DP.
    \item \textbf{Rigorous empirical comparison.} We train a Proximal Policy Optimization agent and compare it against a uniform random policy, a threshold-based heuristic spanning multiple threshold values, and the tabular VI rollout on 100 shared seeds (means, standard deviations, 95\% confidence intervals, Welch tests).
\end{itemize}

\paragraph{Scope.} We deliberately restrict the experiments to a single piece of equipment and a four-dimensional state so that every design choice can be stated, audited and reproduced. This is a term-paper-level study: we aim at providing a clean starting point and an honest baseline for future work on multi-equipment contractual portfolios.

The remainder of the paper is organized as follows. Section~\ref{sec:related_work} reviews related work. Section~\ref{sec:methodology} details the MDP formulation and the learning algorithm. Section~\ref{sec:experiments} describes the experimental protocol and reports the empirical results. Section~\ref{sec:discussion} discusses interpretations, limitations and perspectives, and Section~\ref{sec:conclusion} concludes.