\section{Method}
\label{sec:methodology}

This section formalizes the contract-coverage optimization problem as a Markov decision process (MDP) and describes the reinforcement-learning algorithm we adopt, Proximal Policy Optimization (PPO). We restrict the formulation to the exact scope of our empirical study: a \emph{single piece} of medical equipment followed over a finite daily horizon. This deliberate simplification makes the analysis tractable and keeps the theoretical description and the implementation in strict one-to-one correspondence; extensions to a portfolio of heterogeneous equipment are discussed in Section~\ref{sec:discussion}.


\subsection{MDP formulation}
\label{subsec:mdp}

We model the problem as a finite-horizon MDP $\mathcal{M} = (\mathcal{S}, \mathcal{A}, P, r, T)$, with daily time steps and horizon $T = 120$ days, tracking a single piece of equipment.


\paragraph{State space.}
The state at step $t$ is a four-dimensional vector
\[
    s_t = (\mathrm{covered}_t,\;
           \mathrm{days\_to\_exp}_t,\;
           \mathrm{cost\_level},\;
           \mathrm{risk\_level}),
\]
where $\mathrm{covered}_t \in \{0, 1\}$ indicates whether an active contract covers the equipment, $\mathrm{days\_to\_exp}_t$ is the remaining number of days before expiration of the current coverage, $\mathrm{cost\_level} \in \{0, 1, 2\}$ discretizes the contractual cost (low, medium, high), and $\mathrm{risk\_level} \in \{0, 1, 2\}$ discretizes the equipment risk class (low, medium, high). The two discrete attributes are sampled at the start of each episode and remain constant throughout. All four components are normalized to the unit interval before being fed to the policy.


\paragraph{Action space.}
The action space is discrete with three elements:
\[
    \mathcal{A} = \{\, 0: \text{do nothing},\;
                      1: \text{renew contract},\;
                      2: \text{extend warranty} \,\}.
\]
A \emph{renewal} (action 1) resets the remaining coverage to a fixed duration ($\mathrm{RENEWAL\_DAYS} = 30$) and restores the covered status if the equipment was uncovered. A \emph{warranty extension} (action 2) adds a fixed duration ($\mathrm{EXTENSION\_DAYS} = 30$, capped at $\mathrm{MAX\_COVERED\_DAYS} = 60$) at a reduced cost, but requires an already-active contract; invoking it while uncovered incurs a small no-op penalty.


\paragraph{Reward function.}
The per-step reward aggregates six components, each chosen so that the total remains in a numerically stable range and induces a clear ordering between desirable and undesirable decisions:
\begin{align*}
r_t \;=\; &\; r^{\text{coverage}}_t
          + r^{\text{timing}}_t
          + r^{\text{cost}}_t
          + r^{\text{expiry}}_t
          + r^{\text{early}}_t
          + r^{\text{failure}}_t.
\end{align*}
The continuous coverage term is $+\mathrm{COVERAGE\_BONUS}$ while covered and $-\mathrm{UNCOVERED\_PENALTY}$ otherwise. The timing term rewards renewals or extensions issued close to expiration ($+\mathrm{RENEWAL\_CLOSE\_BONUS}$ when $\mathrm{days\_to\_exp}_t \le 5$) and penalizes early renewals ($-\mathrm{RENEWAL\_EARLY\_PENALTY}$ when $\mathrm{days\_to\_exp}_t \ge 20$). The cost term is a contract cost proportional to $\mathrm{cost\_level}$ (smaller for extensions than for full renewals). The expiration term is a one-shot penalty applied the exact step a contract expires without renewal. The failure term is a one-shot penalty whenever a failure occurs while the equipment is uncovered, sampled with probability proportional to $(1 + \mathrm{risk\_level})$.


\paragraph{Transition dynamics.}
At each step, the environment sequentially (i) applies the chosen action, (ii) decrements $\mathrm{days\_to\_exp}_t$ by one and triggers expiration if the counter reaches zero, (iii) samples a Bernoulli failure indicator with risk-conditional probability, and (iv) emits the scalar reward. The static attributes $\mathrm{cost\_level}$ and $\mathrm{risk\_level}$ are unchanged across a full episode, so the only stochastic transitions concern the coverage counter and the failure process. The dynamics are fully Markovian in the four-dimensional state.


\subsection{Tabular Bellman optimum (value iteration)}
\label{subsec:vi-baseline}

For benchmarking and clarification of optimality notions, we also compute an exact greedy policy in a finite MDP that augments each physical state with the elapsed time index $\tau \in \{0,\dots,T{-}1\}$ recorded by the simulator at each decision.

Because $\mathrm{cost\_level}$ and $\mathrm{risk\_level}$ are fixed throughout an episode, this augmented process is finite---at most $|\mathcal{S}_\text{tab}| \approx 65{,}880$ reachable tuples for our $T{=}120$, $D_{\max}{=}60$ setting---so we solve it with synchronous discounted value iteration using the same $\gamma{=}0{.}99$ as PPO.

The associated greedy policy, denoted \emph{VI (exp.)} in experiments, maximises the usual discounted criterion in \emph{expectation} relative to this tabulation; Monte Carlo evaluation on the sampled simulator inherits Bernoulli noise, so confidence intervals remain non-degenerate.


\subsection{Proximal Policy Optimization (PPO)}
\label{subsec:ppo}

To approximate the MDP defined in Section~\ref{subsec:mdp}, we use the \emph{Proximal Policy Optimization} (PPO) algorithm~\cite{schulman2017ppo}, a policy-gradient method recognized for its empirical stability and implementation simplicity.


\paragraph{Principle.}
PPO belongs to the family of actor-critic methods. The algorithm jointly trains two components: a stochastic policy $\pi_\theta(a \mid s)$ that maps each state to a distribution over the three admissible actions, and a value function $V_\phi(s)$ that estimates the expected future return from state $s$. Both components are parameterized by neural networks sharing a common two-layer encoder (64 units, $\tanh$ activations) with separate softmax policy and scalar value heads. They are learned from trajectories collected under the current policy. The advantage estimate, which measures the marginal gain of an action relative to the average value of the state, is obtained through Generalized Advantage Estimation (GAE)~\cite{schulman2016gae}, with a discount $\gamma = 0.99$ and GAE coefficient $\lambda = 0.95$, providing a controllable trade-off between bias and variance.


\paragraph{Clipped objective.}
The specificity of PPO lies in its learning objective. Let $r_t(\theta) = \pi_\theta(a_t \mid s_t) / \pi_{\theta_{\text{old}}}(a_t \mid s_t)$ denote the likelihood ratio between the current policy and the behavior policy that produced the trajectory, and let $\hat{A}_t$ denote the normalized GAE advantage. PPO maximizes
\[
    \mathcal{L}^{\text{clip}}(\theta)
    = \mathbb{E}_t \Big[
        \min \big(r_t(\theta)\, \hat{A}_t,\;
                 \mathrm{clip}(r_t(\theta),\, 1 - \epsilon,\, 1 + \epsilon)\, \hat{A}_t
            \big)
      \Big],
\]
with clip parameter $\epsilon = 0.2$. The value function is trained by minimizing the squared return-prediction error, and an entropy bonus is added to the global objective to encourage exploration and to prevent premature convergence to a deterministic policy.


\paragraph{Implementation details.}
Observations are normalized online through a running estimate of their mean and variance (Welford's algorithm) and clipped to $[-5, 5]$ before being fed to the network. Advantages are normalized per batch. At each iteration, a full episode is collected under the current policy, and the policy--value pair is updated for four epochs over mini-batches of size $64$. Training is conducted entirely offline on the simulator, which enables intensive exploration without operational risk and yields a reproducible evaluation of the learned policy prior to any deployment.