\section{Related Work}
\label{sec:related_work}

Our study lies at the intersection of three research strands: maintenance and service-contract optimization, reinforcement learning for sequential decision-making in industrial operations, and deep policy-gradient methods.


\paragraph{Maintenance and service-contract optimization.}
Service-contract design and maintenance scheduling have a long tradition in operations research, where they are typically formulated as integer programs, renewal-theoretic models or stochastic dynamic programs~\cite{jardine2006managing, dekker1996applications}. These approaches provide principled solutions under strong modelling assumptions but scale poorly to high-dimensional portfolios with stochastic failures and evolving contractual structures, and typically produce static rather than adaptive policies. In industrial practice, simple threshold rules and segment-based heuristics dominate deployments~\cite{gits1992structuring}, which motivates the search for data-driven alternatives that can capture the temporal feedback between coverage decisions and downstream costs.


\paragraph{Reinforcement learning for industrial operations.}
Reinforcement learning has emerged as a natural framework for sequential decision-making in industrial and financial operations, thanks to its ability to learn from interactions without requiring an explicit model of the environment~\cite{sutton2018reinforcement}. Recent applications include risk-sensitive trading agents trained with language-model-augmented rewards~\cite{benhenda2025}, deep RL for efficient liquidity provisioning in decentralized finance~\cite{xu2025}, and dynamic pricing and inventory management~\cite{boute2022deep}. These works illustrate both the promise of learned decision-making under uncertainty and the recurring difficulties of reward specification, sample efficiency and safe deployment---difficulties that are also central to our setting.


\paragraph{Reinforcement learning in healthcare operations.}
Within healthcare, RL has been applied to treatment recommendation, dynamic treatment regimes and operational decision-making such as ICU ventilator weaning and resource allocation~\cite{yu2021reinforcement}. Most of these applications emphasize safety, interpretability and offline learning from historical logs, all constraints that also apply when managing a fleet of medical devices. To the best of our knowledge, contract-coverage optimization for medical equipment has not yet been cast as an RL problem in the published literature, which motivates the present work.


\paragraph{Policy-gradient methods.}
At the algorithmic level, our approach builds on deep actor-critic methods. Proximal Policy Optimization~\cite{schulman2017ppo} provides an empirically stable policy-gradient procedure through a clipped surrogate objective, which we couple with Generalized Advantage Estimation~\cite{schulman2016gae}. Running observation normalization and advantage normalization, which we also use, are standard components of modern PPO implementations~\cite{andrychowicz2021matters}. Our contribution is not algorithmic but applicative: we design an MDP formulation, a simulation environment and an evaluation protocol tailored to contractual-coverage optimization, and we situate the learned policy against a random baseline and a strong rule-based heuristic.
